Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-03-09 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
hs: additional histograms

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/tmp_corner_plot.pdf}
\caption{(Color online).~Corner plot showing that the inferred final spin $\chi_f$,
detector-frame remnant mass $M_f$, and redshift $z$ for GW150915, using the same waveform model,
but without (blue contours) and with the non-GR parameters different from zero (orange contours, for
dCS gravity and $j=0$). The contours represent 90\% credible regions.
%
We see that the introduction of the non-GR parameters does not bias the inference on the source parameters.}
\label{fig:corner_plot}
\end{figure}

@@

%
\hs{Abhirup, maybe move this to the prior subsection?}
In our parameter estimation runs that we had to adjust our prior on $\ell$ depending on which event and
on which theory we examined.
%
The reasons are twofold.
%
% Describe problem 1
%
First, we found that depending on the source parameters $\btheta$,
large values of $\ell$ could result in unphysical waveforms in the frequency
band we considered. More specifically, these waveforms have a ringdown which is
considerable longer than the inspiral-merger part in the detector frequency band.
%
The fact that our sampled over these waveforms is perhaps not surprising,
because, after all, we are by construction adding corrections to the GR QNM
which \emph{increase} the damping timescale.
%
In practice, we found that these edge-cases were problematic to deal with
\texttt{LALInferenceMCMC}.
%
We overcome this problem, we reduced the prior upper-value of our prior on $\ell$.
%
This choice does not pose a problem to our parameter estimations, because these
waveforms differ grossly from the data and thus have negligible likelihood.
%
% Describe problem 2
%
However, we cannot reduce the prior on $\ell$ too much. If we do so,
to the marginalized posterior distribution on $\ell$ would rail against
the prior or, in a more extreme case, even be a uninformative, flat posterior.
%
% Summary
%
Altogether, this means that the priors $P(\ell)$ are not shared between events and
they must be taken into account when calculation the cumulative posterior function~\eqref{eq:cumulative_dist_ell}.

@@

\hs{Quote 90\% credibility results.}

@@

@@

@@

We presented an unified framework to combine the parametrized formalism introduced
to model deviations to the GR QNMs~\cite{Maselli:2019mjd} with the \pSEOB~waveform model.
%
We showed through concrete examples how theory-specific calculations QNM of
slowly-rotating BHs in modified gravity theories can be mapped upon the
theory-agnostic beyond-GR parameters of the {\sc ParSpec} formalism.
%
Together this allowed us to apply the \pSEOB~waveform model to test three
modified gravity theories (EdGB, dCS and cubic EFTofGR) using real
observational data from the GW events GW150914 and GW200129.
%
We found, in particular that, within the assumptions of our model and by
stacking posteriors, that the coupling constant of dCS gravity is bound to
$\ell_{\rm CS} \leqslant 34.5$~km.
%
In contrast, we could place any constraint of the coupling constants in EdGB gravity $\ell_{\rm GB}$.
%
This dichotomy between these theories has an analogy with works that considered the inspiral part of the GW signal alone.
%
There it is dCS which is unconstrained, while EdGB, in principle can be constrained, based on results for the closely
related shift-symmetric version of this theory.

>>>>

\begin{figure}[t]
\includegraphics[width=\columnwidth]{figs/tmp_corner_plot.pdf}
\caption{(Color online).~Corner plot showing that the inferred final spin $\chi_f$,
detector-frame remnant mass $M_f$, and redshift $z$ for GW150915, using the same waveform model,
but without (blue contours) and with the non-GR parameters different from zero (orange contours, for
dCS gravity and $j=0$). The contours represent 90\% credible regions.
%
We see that the introduction of the non-GR parameters does not bias the inference on the source parameters.
%
\hs{TODO: increase font size and move it to an Appendix.}
}
\label{fig:corner_plot}
\end{figure}

@@

%
\hs{Abhirup, maybe move this to the subsection on priors?}
In our parameter estimation runs that we had to adjust our prior on $\ell$ depending on which event and
on which theory we examined.
%
The reasons are twofold.
%
% Describe problem 1
%
First, we found that depending on the source parameters $\btheta$,
large values of $\ell$ could result in unphysical waveforms in the frequency
band we considered. More specifically, these waveforms have a ringdown which is
considerable longer than the inspiral-merger part in the detector frequency band.
%
The fact that our sampled over these waveforms is perhaps not surprising,
because, after all, we are by construction adding corrections to the GR QNM
which \emph{increase} the damping timescale.
%
In practice, we found that these edge-cases were problematic to deal with
\texttt{LALInferenceMCMC}.
%
We overcome this problem, we reduced the prior upper-value of our prior on $\ell$.
%
This choice does not pose a problem to our parameter estimations, because these
waveforms differ grossly from the data and thus have negligible likelihood.
%
% Describe problem 2
%
However, we cannot reduce the prior on $\ell$ too much. If we do so,
to the marginalized posterior distribution on $\ell$ would rail against
the prior or, in a more extreme case, even be a uninformative, flat posterior.
%
% Summary
%
Altogether, this means that the priors $P(\ell)$ are not shared between events and
they must be taken into account when calculation the cumulative posterior function~\eqref{eq:cumulative_dist_ell}.

@@

\hs{Convention: us 90\% credibility for the final bounds.}

@@

\begin{table}[h]
\begin{tabular}{l l c c c}
\hline
\hline
Event & $j$ & EFT    & ParSpec & Constraint ($\gm = M_{f}$) \\
      &     & bound? & bound?  &                            \\
\hline
GW150914 & $j = 0$    &     &     &    \\
GW150914 & $j = 0, 1$ &     &     &    \\
GW200129 & $j = 0$    &     &     &    \\
GW200129 & $j = 0, 1$ &     &     &    \\
Combined & $j = 0$    &     &     &    \\
Combined & $j = 0, 1$ &     &     &    \\
\hline
\hline
\end{tabular}
\caption{Detailed summary of our results the cubic EFT of GR for GW150914, GW200129, and
combined events using $\gm = M_{f}$, $\varepsilon = 1/2$ and quoting only 90\% credible results.
%
\hs{TODO: update when completed.}
}
\label{tab:summary_ceft}
\end{table}

@@

\begin{table}[h]
\begin{tabular}{l l c c c}
\hline
\hline
Event & $j$ & EFT    & ParSpec & Constraint ($\gm = M_{f}$) \\
      &     & bound? & bound?  &                            \\
\hline
GW150914 & $j = 0$    &     &     &    \\
GW150914 & $j = 0, 1$ &     &     &    \\
GW200129 & $j = 0$    &     &     &    \\
GW200129 & $j = 0, 1$ &     &     &    \\
Combined & $j = 0$    &     &     &    \\
Combined & $j = 0, 1$ &     &     &    \\
\hline
\hline
\end{tabular}
\caption{Detailed summary of our results the quartic EFT of GR for GW150914, GW200129, and
combined events using $\gm = M_{f}$, $\varepsilon = 1/2$ and quoting only 90\% credible results.
%
\hs{TODO: update when completed.}
}
\label{tab:summary_qeft}
\end{table}

@@

We presented an unified framework to combine the parametrized formalism introduced
to model deviations to the GR QNMs~\cite{Maselli:2019mjd} with the \pSEOB~waveform model.
%
We showed through concrete examples how theory-specific calculations QNM of
slowly-rotating BHs in modified gravity theories can be mapped upon the
theory-agnostic beyond-GR parameters of the {\sc ParSpec} formalism.
%
Together this allowed us to apply the \pSEOB~waveform model to test four
modified gravity theories (EdGB, dCS, cubic, and quartic
effective-field-theories of GR) using data from the GW events GW150914 and GW200129.
%
We found, in particular that, within the assumptions of our model and by
stacking posteriors, that the coupling constant of dCS gravity is bound to
$\ell_{\rm CS} \leqslant 34.5$~km.
%
In contrast, we could place any constraint of the coupling constants in EdGB gravity $\ell_{\rm GB}$.
%
This dichotomy between these theories has an analogy with works that considered the inspiral part of the GW signal alone.
%
There it is dCS which is unconstrained, while EdGB, in principle can be constrained, based on results for the closely
related shift-symmetric version of this theory.
